---
title: "p2_expo_David"
author: "Xingxing Tong"
output: html_document
---

- Exploration for Recommend system

###Dataset:u.data ###

```{r}
suppressPackageStartupMessages({
   library(recommenderlab)
   library(dplyr) 
   library(ggplot2) 
   })
```


Basic Info.
1) 100,000 ratings (1-5) from 943 users on 1682 movies
2) Each user has rated at least 20 movies (>=20)
3) Simple demographic info for the users (age, gender, occupation, zip ->"maps" )

Survey: 7-month Sep.19,1997 - April 22,1998 
This data has been cleaned up - users who had less than 20 ratings or did not have complete demographic information were removed from this data set.


Dataset

u.data     -- The full u data set, 100000 ratings by 943 users on 1682 items.
              Each user has rated at least 20 movies.  Users and items are
              numbered consecutively from 1.  The data is randomly
              ordered. This is a tab separated list of 
	         user id | item id | rating | timestamp. 
              The time stamps are unix seconds since 1/1/1970 UTC   



CITATION
==============================================

To acknowledge use of the dataset in publications, please cite the following paper:

F. Maxwell Harper and Joseph A. Konstan. 2015. The MovieLens Datasets:
History and Context. ACM Transactions on Interactive Intelligent
Systems (TiiS) 5, 4, Article 19 (December 2015), 19 pages.
DOI=http://dx.doi.org/10.1145/2827872

```{r}
#setwd("~/GitHub/P2_Recom_Sys_MovieLens/Exploration/David")

# udata<- read.table("http://files.grouplens.org/datasets/movielens/ml-100k/u.data")
# names(udata) <- c("User_id","Item_id","Rating","Timestamp")
# head(udata)

#delete Timestamp
#udata<-udata[,1:3]
# udata <- data.frame(udata)
# save(udata,file="udata.Rda")

# setwd("~/GitHub/P2_Recom_Sys_MovieLens/Exploration/David")

load("./data/udata.Rda")
```


```{r}
Unique_id <-unique(udata$User_id)

id_count<- function(id) {
  return(dim(udata[udata$User_id==id ,])[1])
}
#count how many lines for each unique Id
count_result<-sapply(X=Unique_id,FUN=id_count)
#summary the reuslts

#table(count_result)
#qplot()

count_result_df <- data.frame(count_result)
colnames(count_result_df)[1] <-"Numb_of_Movies"
count_result_df$id <- rownames(count_result_df)

movie_numb_median <- median(count_result_df$Numb_of_Movies)


#ggsave("./image/week7_hist.png",plot=p1)

#divide users into two groups: 
#Movie lovers: rating numb. in [20,65]  Movie enthusiasts: rating numb. in [66,737]

#Movie lovers
u_id_1<-count_result_df[count_result_df$Numb_of_Movies<66,]$id
#Movie enthusiasts
u_id_2<-count_result_df[count_result_df$Numb_of_Movies>65,]$id

```

```{r}
ggplot(count_result_df, aes(Numb_of_Movies)) +geom_histogram(binwidth = 3)+
        geom_vline(xintercept = movie_numb_median,colour="steelblue",linetype = "longdash")+
        geom_vline(xintercept = 100,colour="red",linetype = "longdash")+
        ggtitle("Histogram of movie counts(bw=3)")
```




##Movie

###Exploration of weights estimating



Aim1: Find Top 100 Popular Movies (Popularity depends on # ratings)

Variable: MV_100Id


Aim2: Based on the top 100 movie list find the users who rates those movies.(943 users in total)

Variable:  User_M100


Aim3: For each movie, a list is generated to show which user has rated this movie

Variable:  movie_user_table

Aim4: Estimate weights for each movie based on the information derived above 


```{r}
Unique_Mid <-unique(udata$Item_id)

id_Mcount<- function(id) {
  return(dim(udata[udata$Item_id==id ,])[1])
}

# dim(udata[udata$Item_id==1,])
# id_Mcount(1)


count_Mresult<-sapply(X=Unique_Mid,FUN=id_Mcount)

count_Mresult_df <- data.frame(count_Mresult)
colnames(count_Mresult_df)[1] <-"Numb_of_Ratings"
count_Mresult_df$ItemId <- Unique_Mid

ggplot(count_Mresult_df, aes(Numb_of_Ratings)) +geom_histogram(binwidth = 3)+
        ggtitle("Histogram of Rating counts(bw=3)")


reorder_countM<- count_Mresult_df[order(-count_Mresult_df$Numb_of_Ratings),]

#Top 100 popular Movies
countM_100<-reorder_countM[1:100,]

#delete the rowname
rownames(countM_100) <- NULL

#Range of # Ratings
range(countM_100$Numb_of_Ratings)

#Movies in this list have at least 218 ratings.
ggplot(countM_100, aes(Numb_of_Ratings)) +geom_histogram(bins = 20)+
        ggtitle("Histogram of Rating counts(bin=20)")

#Top 100 popular Movies ID
MV_100Id<-countM_100$ItemId

#People rated 100 popular Movies
User_M100 <- udata[udata$Item_id %in% MV_100Id,]$User_id

#how many people in the list
length(unique(User_M100))


userId_M<-function(Movie_id){
  return( udata[udata$Item_id==Movie_id,]$User_id )
}



# a list of 100 each list indicates user who rated the movie
movie_user_table <-sapply(X=MV_100Id,FUN=userId_M)

# each row indicates a movie
# each column indicates a user
rating_mat <- matrix(data=rep(0,943*100),nrow=100,ncol=943)

for(i in 1:100){
  for(j in 1:943){
    if(j %in% movie_user_table[[i]])
        rating_mat[i,j]=1
  }
}


image(rating_mat)
heatmap(rating_mat)


```
